Inferring the distribution and rate of Earth-like exoplanets \\ from noisy individual-planet inferences

### 2014-04-10 document/ms.tex
note to remember to discuss multiplicity
working on real data

@@

Since the choice of bin locations and number is arbitrary, the most
conservative strategy is to choose very small bins---yielding an effectively
non-parametric model---and choose a prior (or regularization) that encodes any
prior expectations about the distribution.
In particular, it seems reasonable to expect the distribution to be
``smooth''.
The most extreme version of this model would be a completely non-parametric
rate function---generated by, for example, a Gaussian process
(\citealt{poiss-gp})---but this discussion is beyond the scope of this
\paper.
Instead, we'll simply choose small bins and add a Gaussian process prior
(\dfmtodo{CITE}) on the logarithms of the bin heights
\begin{eqnarray}
\ln p(\binheights\,|\,\mu,\,\bvec{\smooth}) &=&
-\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}^{-1}\,\bvec{r}
-\frac{1}{2}\ln \det \bvec{K} - \frac{N}{2}\ln 2\pi
\end{eqnarray}
where $\bvec{r}$ is the vector of residuals
\begin{eqnarray}
\bvec{r} &=& \left(\begin{array}{cccc}
\binheight_1-\mu & \binheight_2-\mu & \cdots & \binheight_N-\mu
\end{array}\right)^\mathrm{T}
\end{eqnarray}
and $\bvec{K}$ is the kernel matrix evaluated \emph{at the bin centers}
$\bincenter_i$
\begin{eqnarray}
K_{ij}(\bvec{\smooth}) &\equiv& \smooth_0^2\,\exp \left(
-\frac{1}{2}[\bincenter_i-\bincenter_j]^\mathrm{T}
\left[\begin{array}{cc}
1/\smooth_1 & 0 \\ 0 & 1/\smooth_2
\end{array}\right]
[\bincenter_i-\bincenter_j] \right) \quad.
\end{eqnarray}

@@

\paragraph{Inverse-detection-efficiency}
Authors in the exoplanet literature (for example, \citealt{howard, dressing,
petigura}) have built population level models using this procedure.
The basic idea is to make a histogram of planet candidates in the model where
each sample has a weight proportional to the inverse detection efficiency.
This method makes intuitive sense and, in the limit of infinitely precise
measurements, it is not wrong.

@@

>>>>

\dfmtodo{%
Cite \citet{tremaine} for likelihood function including multiplicity.
}

@@

Since the choice of bin locations and number is arbitrary, the most
conservative strategy is to choose very small bins---yielding an effectively
non-parametric model---and choose a prior (or regularization) that encodes any
prior expectations about the distribution.
In particular, it seems reasonable to expect the distribution to be
``smooth''.
The most extreme version of this model would be a completely non-parametric
rate function---generated by, for example, a Gaussian process
(\citealt{poiss-gp})---but this discussion is beyond the scope of this
\paper.
Instead, we'll simply choose small bins and add a Gaussian process prior
(\dfmtodo{CITE}) on the logarithms of the bin heights
\begin{eqnarray}
\ln p(\binheights\,|\,\mu,\,\bvec{\smooth}) &=&
-\frac{1}{2}\,\bvec{r}^\mathrm{T}\,\bvec{K}^{-1}\,\bvec{r}
-\frac{1}{2}\ln \det \bvec{K} - \frac{N}{2}\ln 2\pi
\end{eqnarray}
where $\bvec{r}$ is the vector of residuals
\begin{eqnarray}
\bvec{r} &=& \left(\begin{array}{cccc}
\binheight_1-\mu & \binheight_2-\mu & \cdots & \binheight_N-\mu
\end{array}\right)^\mathrm{T}
\end{eqnarray}
and $\bvec{K}$ is the kernel matrix evaluated \emph{at the bin centers}
$\bincenter_i$
\begin{eqnarray}
K_{ij}(\bvec{\smooth}) &\equiv& \smooth_0^2\,\exp \left(
-\frac{1}{2}[\bincenter_i-\bincenter_j]^\mathrm{T}\,
\mathrm{diag}(1/\smooth_k^2)\,
[\bincenter_i-\bincenter_j] \right) \quad.
\end{eqnarray}

@@

\paragraph{Inverse-detection-efficiency}
Authors in the exoplanet literature (for example, \citealt{howard, dressing,
petigura}) have built population level models using this procedure.
The basic idea is to make a histogram of planet candidates in the model where
each sample has a weight proportional to the inverse detection efficiency.
This method makes intuitive sense and, in the limit of infinitely precise
measurements, it is not wrong.
Even in this limit, this estimator of the occurrence distributions will have
large variance; especially as the detection efficiency becomes low.
In the presence of observational uncertainties, the inferred distributions
will be biased.

@@

\bibitem[Tremaine \& Dong(2012)]{tremaine}
Tremaine, S., \& Dong, S.\ 2012, \aj, 143, 94 (\arxiv{1106.5403})
